A Novel Model-Free Learning Approach for Dynamic Target Tracking Problem and its Validation using Virtual Robot Experimentation Platform

### 2020-07-02 paperMIC2020/paperIEEE-MIC2020-V1.tex
Finished RL section

 The critic weights are then used to update the actor weights $W_a\mathbb{R}^{n\times m}$ (here $n=2$ and $m=2$) which maps the state error to the desired policy (control actions) as  at every time step $t$ with the following equation:
 \begin{align}
 \label{eq:actorWeightsPolicy}    
 \mathbf{u}_k  =  {\bf W}_a^T \mathbf{e}_k
 \end{align}
%
 Equation~\eqref{eq:actorWeightsPolicy} is realized using a multi-layer actor neural network shown in Fig.~

 Applying a similar gradient descent approach as with the critic weights the actor weights are updated as follows:
\begin{align}
   \bm{\omega_a}^{[r+1]} = \bm{\omega}^{[r]} - \ell_a(\mathbf{u}_k-\mathbf{u}_k^*)\mathbf{e}_k, 
 \label{eq:actorWeights}
 \end{align}
 where $\ell$ is the actor learning rate. 
 \\
 The complete policy iteration solution process for the leader-follower problem is detailed out in Algorithm \ref{alg:ModelFreeTracking}. %
 %
 \begin{algorithm2e}{
     \caption{\label{alg:ModelFreeTracking} Model-free actor-critic reinforcement learning using the policy iteration solution.}
     \DontPrintSemicolon
     \KwIn{Sampling-time $T_s,$ $\mathbf{Q},~\text{and}~\mathbf{R}$ }
     \KwOut{Error trajectory $\mathbf{e}_k,$ for $k=0,1,\ldots$}
     % \KwData{Map representation using occupancy grid technique}
     % \KwResult{output******************************}
     \Begin{
         $k=0, r = 0$ \tcc*[h]{Discrete time and policy indices}\;
         $\eta = (n+m)(n+m+1)/2$\;
         Initialize $\mathbf{P}^{[0]}$ \tcc*[h]{Positive definite}\;
                 Set offset distance $d$\;
                 Given approximate initial poses of leader and follower, compute $\mathbf{e}_0$ using error model~\eqref{eq:stateError}\;
                 Compute follower's input $\mathbf{u}_0^{[0]}$ using policy~\eqref{eq:actorWeightsPolicy}\;
         \Repeat(\tcc*[h]{Main timing loop}){Tracking errors are zero}
         {
             Find $\mathbf{e}_{k+1}$ using~\eqref{eq:stateError}\;
             Compute policy $\mathbf{u}_{k+1}^{[r]}$ using~\eqref{eq:actorWeightsPolicy}\;
             %Evaluate and record Equation~\eqref{eq:modelFreeMatrixSolution}\;
             \If{[$(k+1)~\mathrm{modulo}~\eta]==0$ }
             {
                 $r\leftarrow r+1$\tcc*[h]{Evaluate policy}using~\eqref{eq:modelFreePolicy}\;
                 Solve for the critic-weights $\bm{\omega_c}$ using~\eqref{eq:criticWeights}eights}\;
                 Solve for the actor-weights $\bm{\omega_a}$ using ~\eqref{eq:actorWeights}\;\;
                 Construct matrix $\mathbf{P}^{[r]}$ using vector $\bm{\omega_c}$\;
 %                \eIf{$\|\mathbf{P}^{[r]} - \mathbf{P}^{[r+1]}\|<\varepsilon$}{Set $\mathbf{u}_{k+1}^*\leftarrow\mathbf{u}_{k+1}^{[r]}$}{$k\leftarrow k+1$ }
             }
 %        {
 %                $k\leftarrow k+1$
 %            }
         $k\leftarrow k+1$
         }
     }
 \end{algorithm2e}

>>>>

 The critic weights are then used to update the actor weights $W_a\mathbb{R}^{n\times m}$ (here $n=2$ and $m=2$) which maps the state error to the desired policy (control actions) as  at every time step $t$ with the following equation:
 \begin{align}
 \label{eq:actorWeightsPolicy}    
 \hat{\mathbf{u}}_k  =  {\bf W}_a^T \bm{\sigma}(\mathbf{e}_k),
 \end{align}
 %
 where ${\bf W}_a\in\mathbb{R}^{2\times n^{'}}$ is the actor weight matrix and $\bm{\sigma}:\mathbb{R}^2 \to \mathbb{R}^{n^{'}}$ is the basis function vector.  Equation~\eqref{eq:actorWeightsPolicy} is realized using a multi-layer actor neural network shown in Fig.~\ref{fig:nnActor}. %
 %
\begin{figure}
  \centering
  \boxed{
    \begin{adjustbox}{max width = 0.48\textwidth}
      \tikzset{%
        input neuron/.style={
          circle,
          fill=green!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

        \foreach \m/\l [count=\y] in {1,2}
        \node [input neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

        \foreach \m [count=\y] in {1,2,missing,3}
        \node [hidden neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2.5-\y*1.1) {};

        \foreach \m [count=\y] in {1,2}
        \node [output neuron/.try, neuron \m/.try] (output-\m) at (4,2.5-\y*1.2) {};

        \foreach \l [count=\i] in {1,2}
        \draw [<-] (input-\i) -- ++(-1,0)
        node [above, midway] {$\mathbf{e}_{k,\l}$};

        \foreach \l [count=\i] in {1,2,\mbox{$n^{'}$}}
        \node [above] at (hidden-\i.north) {$\sigma_{\l}$};

        \foreach \l [count=\i] in {1,2}
        \draw [->] (output-\i) -- ++(1,0)
        node [above, midway] {$\hat{u}_{k,\l}$};

        \foreach \i in {1,...,2}
        \foreach \j in {1,...,3}
        \draw [->] (input-\i) -- (hidden-\j);

        \foreach \i in {1,...,3}
        \foreach \j in {1,2}
        \draw [->] (hidden-\i) -- (output-\j);

        \foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
        \node [align=center, above] at (\x*2,1.9) {\l \\ layer};

        \draw[->,ultra thick]
        (3,-1.5)node[anchor=west]{Weight~${\bf W}_a$} to[bend left] (3,1.6);
  \caption{Actor neural network structure for approximating control input.}
  \label{fig:nnActor}
The elements of the vector $\bm{\sigma}$ are the activation functions of the neurons in the hidden layer shown in Fig.~\ref{fig:nnActor}. We will need to determine the actor weight matrix ${\bf W}_a$ that minimizes the error between the estimated control inputs determined by model~\eqref{eq:actorWeightsPolicy} and the that determined by critic neural network approximation model~\eqref{eq:modelFreePolicy}. For that, let us define the error %
  \label{eq:deltaActor}
  \delta_a = \frac{1}{2}\|\hat{{\bf u}}_k - {\bf u}_k^*\|^2,
which can be written as %
  \delta_a = \frac{1}{2}\left( {\bf W}_a^T\bm{\sigma}({\bf e}_k) -  {\bf u}_k^*\right)^T\left( {\bf W}_a^T\bm{\sigma}({\bf e}_k) -  {\bf u}_k^*\right). 
Taking partial derivative of the above expression $\delta_a$  with respect to ${\bf W}_a$ gives %
  \frac{\partial \delta_a}{\partial{\bf W}_a} = {\bf W}_a^T\bm{\sigma}({\bf e}_k)\bm{\sigma}^T({\bf e}_k) - {\bf u}_k^*\bm{\sigma}^T({\bf e}_k).
Therefore, actor weight matrix ${\bf W}_a$ in every policy iteration $r = 01,2,\ldots$ can be updated using the learning rule %
  {\bf W}_a^{[r+1]} = {\bf W}_a^{[r]} - \ell_a\frac{\partial\delta_a}{\partial{\bf W}_a}\\
  = {\bf W}_a^{[r]} - \ell_a\left( {\bf W}_a^T\bm{\sigma}({\bf e}_k)-\mathbf{u}_k^* \right)\bm{\sigma}^T({\bf e}_k), 
 \label{eq:actorWeights}
 where $\ell_a$ is the actor learning rate.  The complete policy iteration solution process for the leader-follower problem is detailed out in Algorithm \ref{alg:ModelFreeTracking}. %
 \begin{algorithm2e}{
     \caption{\label{alg:ModelFreeTracking} Model-free actor-critic reinforcement learning using the policy iteration solution.}
     \DontPrintSemicolon
     \KwIn{Sampling-time $T_s,$ $\mathbf{Q},~\text{and}~\mathbf{R}$ }
     \KwOut{Error trajectory $\mathbf{e}_k,$ for $k=0,1,\ldots$}
     % \KwData{Map representation using occupancy grid technique}
     % \KwResult{output******************************}
     \Begin{
         $k=0, r = 0$ \tcc*[h]{Discrete time and policy indices}\;
         $\eta = (n+m)(n+m+1)/2$\;
         Initialize $\mathbf{P}^{[0]}$ \tcc*[h]{Positive definite}\;
                 Set offset distance $d$\;
                 Given approximate initial poses of leader and follower, compute $\mathbf{e}_0$ using error model~\eqref{eq:stateError}\;
                 Compute follower's input $\mathbf{u}_0^{[0]}$ using policy~\eqref{eq:actorWeightsPolicy}\;
         \Repeat(\tcc*[h]{Main timing loop}){Tracking errors are zero}
         {
             Find $\mathbf{e}_{k+1}$ using~\eqref{eq:stateError}\;
             Compute policy $\mathbf{u}_{k+1}^{[r]}$ using~\eqref{eq:actorWeightsPolicy}\;
             %Evaluate and record Equation~\eqref{eq:modelFreeMatrixSolution}\;
             \If{[$(k+1)~\mathrm{modulo}~\eta]==0$ }
             {
                 $r\leftarrow r+1$\tcc*[h]{Evaluate policy}using~\eqref{eq:modelFreePolicy}\;
                 Solve for the critic-weights $\bm{\omega_c}$ using~\eqref{eq:criticWeights}eights}\;
                 Solve for the actor-weights $\bm{\omega_a}$ using ~\eqref{eq:actorWeights}\;\;
                 Construct matrix $\mathbf{P}^{[r]}$ using vector $\bm{\omega_c}$\;
 %                \eIf{$\|\mathbf{P}^{[r]} - \mathbf{P}^{[r+1]}\|<\varepsilon$}{Set $\mathbf{u}_{k+1}^*\leftarrow\mathbf{u}_{k+1}^{[r]}$}{$k\leftarrow k+1$ }
 %        {
 %                $k\leftarrow k+1$
 %            }
         $k\leftarrow k+1$
 \end{algorithm2e}    
For clarity, number of hidden neurons chosen in the actor neural network shown in Fig.~\ref{fig:nnActor} is $n^{'} = 2$ and the basis function vector is simply $\bm{\sigma}({\bf e}_k) = {\bf e}_k.$ In the following, we provide the computer simulation results to highlight the performance of the proposed model-free trajectory tracking approach for a mobile robot in tracking a random target while maintaining a geometric safe distance $d>0.$
